Parallel Monte Carlo Synthetic Acceleration Methods for Discrete Transport Problems

### 2013-08-28 doc/dissertation/presentation/defense.tex
updating defense presentation

  \begin{itemize}
  \item Modern hardware is moving in two directions (Kogge,2011):
    \begin{itemize}
    \item Lightweight machines
    \item Heterogeneous machines
    \item Both characterized by low power and high concurrency
    \end{itemize}
    \medskip \medskip
  \item Some issues:
    \begin{itemize}
    \item Higher potential for both soft and hard failures (DOE,2012)
    \item Memory restrictions are expected with a continued decrease
      in memory/FLOPS
    \end{itemize}
    \medskip \medskip
  \item Potential resolution from Monte Carlo:
    \begin{itemize}
    \item Soft failures buried within the tally variance
    \item Hard failures are high variance events
    \item Memory savings over conventional methods
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item Physics-driven development
    \begin{itemize}
    \item Identify the physics of interest and research MCSA
      applicability
    \item Offer solutions or a potential path forward to the observed
      issues
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear scheme for discrete systems leveraging
    Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to study MCSA preconditioning
    \item Iterative performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear scheme for discrete systems
    leveraging Monte Carlo Synthetic Acceleration
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Parallel scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Jacobi Preconditioned Calculations}

@@

  \begin{itemize}
  \item Sequence of Navier-Stokes benchmarks for the nonlinear methods
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
    \medskip
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
    \medskip
  \end{itemize}

  \medskip  \medskip
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \[
      \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho
      \ve{g} = \ve{0}
      \]
      \[
      \nabla \cdot \ve{u} = 0
      \]
      \[
      \rho C_p \ve{u} \cdot \nabla T + \nabla \cdot \ve{q} = 0
      \]
    \end{column}

    \begin{column}{0.5\textwidth}
      \[
      \ve{T} = -P \ve{I} + \mu[\nabla \ve{u} + \nabla \ve{u}^T]
      \]
      \[
      \ve{q} = - k \nabla T
      \]
    \end{column}
  \end{columns}

@@

%%----------s-----------------------------------------------------------------%%
\begin{frame}{MCSA for Nonlinear Problems}

  \begin{itemize}
  \item Need a nonlinear solution scheme for the Navier-Stokes equations
    \bigskip
  \item Signficant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Potential parallel scaling and iterative performance
      improvements
    \item Potential resiliency benefits
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
    {\small
    \item Drekar is a production fluid code base using Newton methods
      with FAD
      \medskip
    \item MCLS incorporated into Drekar nonlinear scheme to implement
      FANM
      \medskip
    \item Newton-Krylov method with Aztec GMRES used for benchmark
      comparisons
    \item All problems preconditioned with explicit scheme using
      algebraic multigrid (ML) and leveraged some kind of
      globalization (e.g. backtracking) 
    }
  \end{itemize}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_pure_strong_time.pdf}
        \end{center}
        \caption{CPU Time.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_pure_strong.pdf}
        \end{center}
        \caption{Absolute Efficiency.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_pure_weak_time.pdf}
        \end{center}
        \caption{CPU Time.}
      \end{figure}

@@

    \begin{figure}[htpb!]
      \begin{center}
        \includegraphics[width=2.4in]{titan_weak_absolute.pdf}
      \end{center}
      \caption{Absolute Efficiency.}
    \end{figure}

@@

  \begin{itemize}
  \item Improved implementation efficiency means improved
    iterative performance will potentially give a better time to
    solution
    \bigskip
  \item Implementation efficiency is effectively the parallel
    efficiency of a single iteration
  \end{itemize}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_ms_eff.pdf}
        \end{center}
        \caption{\small{Absolute parallel efficiency relative to
            16-core 1-set base case.}}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_ms_time.pdf}
        \end{center}
        \caption{\small{Wall time in seconds to solution for each
            case.}}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_weak_ms_eff.pdf}
        \end{center}
        \caption{Absolute parallel efficiency relative to 16-core
          1-set base case.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_weak_ms_time.pdf}
        \end{center}
        \caption{Wall time in seconds to solution for each case.}
      \end{figure}

@@

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_subdomain.pdf}
        \end{center}
        \caption{Strong scaling absolute efficiency for pure domain
          decomposition.}
        \label{fig:titan_strong_subdomain}
      \end{figure}

@@

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_weak_subdomain.pdf}
        \end{center}
        \caption{Weak scaling absolute efficiency for pure domain
          decomposition.}
        \label{fig:titan_weak_subdomain}
      \end{figure}

>>>>

  \begin{itemize}
  \item Modern hardware is moving in two directions (Kogge,2011):
    \begin{itemize}
    \item Lightweight machines
    \item Heterogeneous machines
    \item Both characterized by low power and high concurrency
    \end{itemize}
    \medskip \medskip
  \item Some issues:
    \begin{itemize}
    \item Higher potential for both soft and hard failures (DOE,2012)
    \item Memory restrictions are expected with a continued decrease
      in memory/FLOPS
    \end{itemize}
    \medskip \medskip
  \item Potential resolution from Monte Carlo:
    \begin{itemize}
    \item Soft failures buried within the tally variance
    \item Hard failures mitigated by replication
    \item Memory savings over conventional methods
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
  \item New algorithms required to leverage new computational
    resources
    \begin{itemize}
    \item Tremendous computational resources are required with
      $O(\sn{1}{9})$ element meshes and $O(100,000)+$ cores used today
      for neutronics and fluid problems (Evans,2010)(Pawlowski,2012)
    \end{itemize}
    \bigskip
  \item Necessary to determine applicability to reactor physics and
    potential performance improvements before moving forward with
    other work
    \bigskip
  \item Physics-driven development
    \begin{itemize}
    \item Research applicability and potential improvements to
      neutronics and fluid flow
    \item Offer solutions or a potential path forward for the observed
      issues
    \item Work to improve iterative and parallel performance where
      possible
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Research Outline}
  \begin{itemize}
  \item Development of a linear scheme for discrete systems leveraging
    Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Application to neutron transport
    \item Research is required to study MCSA preconditioning
    \item Iterative performance is of concern
    \end{itemize}
    \bigskip
  \item Development of a nonlinear scheme for discrete systems
    leveraging Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Monte Carlo is a more natural fit
    \item Application to fluid flow
    \item Convergence of the linear model is of concern
    \end{itemize}
    \bigskip
  \item Parallelization of Monte Carlo Synthetic Acceleration
    \medskip
    \begin{itemize}
    \item Parallel strategies taken from modern reactor physics
      methods
    \item Research is required to explore varying parallel strategies
    \item Parallel scalability is of concern
    \end{itemize}
  \end{itemize}
\end{frame}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Block Jacobi Preconditioned Calculations}

@@

  \begin{itemize}
  \item Sequence of Navier-Stokes benchmarks
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
    \medskip
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
    \medskip
  \end{itemize}

  \begin{columns}
    \begin{column}{0.4\textwidth}
      {\small
      \[
      \rho \ve{u} \cdot \nabla \ve{u} - \nabla \cdot \ve{T} - \rho
      \ve{g} = \ve{0}
      \]
      \[
      \nabla \cdot \ve{u} = 0
      \]
      \[
      \rho C_p \ve{u} \cdot \nabla T + \nabla \cdot \ve{q} = 0
      \]
      \[
      \ve{T} = -P \ve{I} + \mu[\nabla \ve{u} + \nabla \ve{u}^T]
      \]
      \[
      \ve{q} = - k \nabla T
      \]
      }
    \end{column}

    \begin{column}{0.6\textwidth}

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.9in]{benchmark_solutions.png}
        \end{center}
      \end{figure}

@@

%%----------s-----------------------------------------------------------------%%
\begin{frame}{MCSA for the Navier-Stokes Equations}

  \begin{itemize}
  \item Need a nonlinear solution scheme for the Navier-Stokes equations
    \bigskip
  \item Significant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Similar framework properties to matrix-free methods
    \end{itemize}
  \end{itemize}

@@

  \begin{itemize}
    {\small
    \item Drekar is a production fluid base using Newton methods
      with FAD
      \medskip
    \item MCLS incorporated into Drekar nonlinear scheme to implement
      FANM
      \medskip
    \item Newton-Krylov method with Aztec GMRES used for benchmark
      comparisons
    \item All problems preconditioned with explicit scheme using
      algebraic multigrid (ML) and leveraged some kind of
      globalization (e.g. backtracking) 
    }
  \end{itemize}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_pure_strong_time.pdf}
        \end{center}
        \caption{Wall time.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_pure_strong.pdf}
        \end{center}
        \caption{Absolute efficiency.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.3in]{titan_pure_weak_time.pdf}
        \end{center}
%        \caption{Wall Time.}
      \end{figure}

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.3in]{titan_weak_implementation.pdf}
        \end{center}
%        \caption{Implementation Efficiency.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.3in]{titan_weak_absolute.pdf}
        \end{center}
%        \caption{Absolute Efficiency.}
      \end{figure}

      {\small
        \begin{itemize}
        \item Krylov implementation efficiencies drop with
          reduced iteration count
        \item MCSA maintained constant number of iterations
        \end{itemize}
      }

@@

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_ms_eff.pdf}
        \end{center}
        \caption{\small{Absolute efficiency relative to 16-core 1-set
            base case.}}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_ms_time.pdf}
        \end{center}
        \caption{\small{Wall time}}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_weak_ms_eff.pdf}
        \end{center}
        \caption{Absolute efficiency relative to 16-core 1-set base
          case.}
      \end{figure}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.4in]{titan_weak_ms_time.pdf}
        \end{center}
        \caption{Wall time.}
      \end{figure}

@@

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_strong_subdomain.pdf}
        \end{center}
        \caption{Strong scaling absolute efficiency relative to full
          transport base case.}
        \label{fig:titan_strong_subdomain}
      \end{figure}

@@

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=2.35in]{titan_weak_subdomain.pdf}
        \end{center}
        \caption{Weak scaling absolute efficiency relative to full
          transport base case.}
        \label{fig:titan_weak_subdomain}
      \end{figure}
